---
title: "Getting Started with scimagR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with scimagR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

scimagR provides an end-to-end pipeline for longitudinal MRI and CT analysis
in spinal cord injury (SCI) research. This vignette walks you through:

1. Installing prerequisites
2. Creating a new project
3. Organizing your data
4. Running the pipeline
5. Analyzing results

## Prerequisites

scimagR requires the following external tools:

- **Spinal Cord Toolbox (SCT)** >= 6.4: <https://spinalcordtoolbox.com>
- **dcm2niix**: <https://github.com/rordenlab/dcm2niix>
- **Python 3** with pydicom: `pip install pydicom`

Check that all tools are available:

```{r check-tools}
library(scimagR)

check_sct()
check_dcm2niix()
check_pydicom()
```

## Creating a Project

Use `sci_create_project()` to scaffold a new workflowr-based analysis project:
```{r create-project}
sci_create_project(
  path = "~/projects/my-sci-study",
  title = "Cervical SCI Longitudinal MRI Analysis",
  author = "Your Name"
)
```

This creates a complete directory structure with template files, configuration,
and analysis Rmd notebooks.

## Data Organization

Place your DICOM files in `data/raw/dicom/` following this structure:

```
data/raw/dicom/
тФЬтФАтФА SCI001/
тФВ   тФЬтФАтФА ses-01/
тФВ   тФВ   тФФтФАтФА *.dcm
тФВ   тФФтФАтФА ses-02/
тФВ       тФФтФАтФА *.dcm
тФФтФАтФА SCI002/
    тФФтФАтФА ses-01/
        тФФтФАтФА *.dcm
```

Fill in the CSV templates in `data/metadata/`:

- `imaging_registry.csv`: One row per imaging session
- `clinical_data.csv`: One row per patient

## Running the Pipeline

The pipeline orchestrates all processing steps:

```{r run-pipeline}
sci_run_pipeline("~/projects/my-sci-study", steps = 1:8)
```

Check pipeline status at any time:

```{r pipeline-status}
sci_pipeline_status("~/projects/my-sci-study")
```

## Computing SCI Metrics

scimagR includes validated implementations of common SCI metrics:

```{r metrics}
# Maximum Spinal Cord Compression
compute_mscc(di = 5.2, da = 8.1, db = 8.5)

# Compression ratio
compute_compression_ratio(ap = 6.5, transverse = 12.0)

# Phase classification
classify_phase(c(0, 5, 30, 200, 400))
```

## Visualization

Use the built-in theme and palettes for publication-ready figures:

```{r viz}
library(ggplot2)

ggplot(data, aes(x = phase, y = mscc, fill = phase)) +
  geom_violin() +
  scale_fill_phase() +
  theme_sci()
```
