---
title: "Pipeline Walkthrough"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pipeline Walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

`sci_run_pipeline()` orchestrates the full SCI imaging processing workflow
in 8 sequential steps. Each step is idempotent and the pipeline can be
interrupted and resumed at any point.

## Pipeline Steps

| Step | Name | Tool | Output |
|------|------|------|--------|
| 1 | DICOM metadata extraction | pydicom | metadata CSV |
| 2 | Anonymization (optional) | pydicom | anonymized DICOMs + mapping |
| 3 | DICOM to NIfTI conversion | dcm2niix | .nii.gz files |
| 4 | Spinal cord segmentation | SCT | _seg.nii.gz |
| 5 | Lesion segmentation | SCT (SCIsegV2) | _lesion_seg.nii.gz |
| 6 | Vertebral labeling | SCT | _labels-disc.nii.gz |
| 7 | Parameter extraction | SCT | CSVs with metrics |
| 8 | Merge & export | R | combined dataset |

## Running the Full Pipeline

```{r full-pipeline}
library(scimagR)

status <- sci_run_pipeline(
  project_dir = "~/projects/my-sci-study",
  steps = 1:8,
  anonymize = FALSE
)
```

## Running Individual Steps

Run only specific steps by setting the `steps` argument:

```{r partial}
# Just convert DICOMs
sci_run_pipeline("~/projects/my-sci-study", steps = 3)

# Run segmentation steps
sci_run_pipeline("~/projects/my-sci-study", steps = 4:5)
```

## The QC Pause

After Step 5 (lesion segmentation), the pipeline pauses for manual QC.
This is critical: automated lesion segmentation must be visually verified
before extracting quantitative parameters.

During the QC pause:

1. Open the QC HTML reports in `data/qc/`
2. Grade each segmentation in the QC log
3. Re-run failed segmentations if needed
4. Continue with steps 6-8

## Checking Pipeline Status

```{r status}
sci_pipeline_status("~/projects/my-sci-study")
```

This returns a tibble showing which steps are complete for each
patient and session.

## Artifact Grading Strategy

After visual QC, assign artifact grades (0-4) in the imaging registry:

- **0 (none)**: Perfect image quality
- **1 (mild)**: Minor artifacts, measurements reliable
- **2 (moderate)**: Visible artifacts, measurements acceptable
- **3 (severe)**: Major artifacts, measurements unreliable
- **4 (non-evaluable)**: Cannot be analyzed

Use `filter_evaluable()` to exclude sessions with grade > 2:

```{r filter}
evaluable <- filter_evaluable(registry, max_artifact = 2)
```

## Coverage Matrix

After processing, check your data coverage:

```{r coverage}
plot_coverage(registry)
```

This tile plot shows which modalities are available for each patient
at each timepoint, with artifact grades overlaid.

## Resuming After Interruption

The pipeline is designed to be resilient to interruptions. Simply re-run
with the same arguments — existing outputs will be detected and skipped:

```{r resume}
# Re-running skips completed steps automatically
sci_run_pipeline("~/projects/my-sci-study", steps = 1:8)

# Force re-processing with overwrite = TRUE
sci_run_pipeline("~/projects/my-sci-study", steps = 4:5, overwrite = TRUE)
```
